---
title: "R Notebook"
output: html_notebook
---

## Part 1: Statistical Analysis
For the first part of this project, you'll mimic the methodology of the working paper ["Does Federally Subsidized Rental Housing Depress Neighborhood Property Values?"](https://furmancenter.org/research/publication/does-federally-subsidized-rental-housing-depress-neighborhood-property), building a statistical model to explore the effect on sales price of a home being within a half mile of an affordable housing development.

1. Using the sf library, find the closest development to each home.

```{r}
library(dplyr)

library(ggplot2)

library(tidyverse)

library(sf)

```
```{r}
#bringing the data in
sales  <-  read.csv('data/filtered_sales.csv')

details  <-  read.csv('data/property_details.csv')

lihtc  <-  read.csv('data/LIHTC.csv')

barnes  <-  read.csv('data/barnes.csv')
```
```{r}
sale_det <- left_join(sales, details, on = apn) %>%
  mutate(centroid = str_remove_all(centroid, '[()]')) %>% 
  separate(centroid, into = c('lon', 'lat'), sep = ',')  
              
sale_det

```
```{r}
sale_det_sf <- st_as_sf(sale_det, coords = c("lon", "lat"))
glimpse(sale_det_sf)
```

```{r}
#need to look at barnes and lihtc and figure out how to combine key components
```

